---
title: "p8105_hw6_zz3166"
author: "Zihan Zhao"
date: "2024-12-01"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(broom)
library(purrr)
library(modelr)
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```


# Problem 2

## Introduction
This report analyzes homicide data from 50 large U.S. cities to understand the factors associated with solving homicides.

### Data Preparation and Cleaning
```{r}
# Load the dataset
homicide_data <- read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv")

# Data cleaning
homicide_data_clean <- homicide_data %>%
  mutate(
    city_state = paste(city, state, sep = ", "),
    solved_binary = ifelse(disposition == "Closed by arrest", 1, 0),
    victim_age = as.numeric(victim_age)
  ) %>%
  filter(
    !(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")),
    victim_race %in% c("White", "Black"),
    !is.na(victim_age)
  )
```

### Logistic Regression: Baltimore, MD
For Baltimore, MD, we fit a logistic regression model with `solved_binary` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors.
```{r}
# Filter for Baltimore
baltimore_data <- homicide_data_clean %>% filter(city_state == "Baltimore, MD")

# Fit logistic regression
baltimore_glm <- glm(
  solved_binary ~ victim_age + victim_sex + victim_race,
  data = baltimore_data,
  family = binomial
)

# Extract adjusted ORs and CIs
baltimore_results <- broom::tidy(baltimore_glm, exponentiate = TRUE, conf.int = TRUE) %>%
  filter(term == "victim_sexMale") %>%
  select(term, estimate, conf.low, conf.high)

# Display results
baltimore_results
```


The adjusted odds ratio for solving homicides involving male victims compared to female victims in Baltimore 

- Odds Ratio (Estimate): 0.425
- 95% Confidence Interval: [0.324, 0.558]

This indicates that male victims are less likely to have their homicides solved compared to female victims, keeping other variables fixed.


### Logistic Regression for All Cities
We repeat the logistic regression for each city and extract the adjusted odds ratios and confidence intervals for solving homicides comparing male to female victims.

```{r}
# Group data by city and run logistic regression for each
city_results <- homicide_data_clean %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    model = map(data, ~ glm(
      solved_binary ~ victim_age + victim_sex + victim_race,
      data = .x,
      family = binomial
    )),
    results = map(model, ~ broom::tidy(.x, exponentiate = TRUE, conf.int = TRUE))
  ) %>%
  unnest(results) %>%
  filter(term == "victim_sexMale") %>%
  select(city_state, term, estimate, conf.low, conf.high)

print(city_results)

```

### Visualization: Adjusted ORs Across Cities
We plot the adjusted odds ratios and confidence intervals for solving homicides, organized by city.
```{r}
# Plot ORs and CIs
city_results %>%
  arrange(estimate) %>%
  mutate(city_state = fct_reorder(city_state, estimate)) %>%
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  coord_flip() +
  labs(
    title = "Adjusted Odds Ratios for Solving Homicides by City",
    x = "City",
    y = "Adjusted Odds Ratio (Male vs Female Victims)"
  )

```

The plot shows significant variability in adjusted odds ratios (ORs) for solving homicides across cities, with most ORs below 1, indicating male victims are less likely to have their homicides solved compared to female victims. Cities like Baltimore, MD, and Chicago, IL, have particularly low ORs, highlighting disparities, while others like Albuquerque, NM, have ORs above 1, suggesting the opposite trend. Wide confidence intervals for some cities reflect uncertainty due to smaller sample sizes, whereas narrow intervals in cities like Baltimore indicate more robust estimates. These disparities underscore the need to investigate systemic factors and practices that contribute to such differences in homicide resolution.



# Problem 3

## Introduction
The dataset contains information on roughly 4000 children, capturing various demographic, physical, and behavioral factors related to their birthweight. This study investigates the relationship between these factors and birthweight using regression models. A proposed model is compared with two alternative models, with evaluation based on residual analysis and cross-validation error.
```{r}
# Load the data
birthweight_raw <- read_csv("birthweight.csv")

# Clean and preprocess the dataset
birthweight_clean <- birthweight_raw |>
  mutate(
    # Recode categorical variables
    mrace = case_match(mrace, 
      1 ~ "White", 
      2 ~ "Black", 
      3 ~ "Asian", 
      4 ~ "Puerto Rican", 
      8 ~ "Other"),
    frace = case_match(frace, 
      1 ~ "White", 
      2 ~ "Black", 
      3 ~ "Asian", 
      4 ~ "Puerto Rican", 
      8 ~ "Other", 
      9 ~ "Unknown"),
    babysex = case_match(babysex, 
      1 ~ "Male", 
      2 ~ "Female"),
    blength = as.numeric(blength),
    bhead = as.numeric(bhead),
    gaweeks = as.numeric(gaweeks),
    ppwt = as.numeric(ppwt),
    wtgain = as.numeric(wtgain)
  ) |>
  drop_na()

# Check the structure of the cleaned dataset
summary(birthweight_clean)
```
The dataset contains 4342 observations with 20 variables.

## Proposed Regression Model

The first model hypothesizes that birthweight is influenced by *gestational age* (`gaweeks`), *maternal pre-pregnancy weight* (`ppwt`), and *birth length* (`blength`), as well as an interaction between `gaweeks` and `blength`.

```{r}
# Fit the proposed model
proposed_model <- lm(bwt ~ babysex + gaweeks + ppwt + blength + blength:gaweeks, data = birthweight_clean)

# Summarize the proposed model
proposed_model_summary <- proposed_model |> broom::tidy()
proposed_model_summary |>
  mutate(
    OR = exp(estimate), 
    CI_upper = exp(estimate + 1.96 * std.error),
    CI_lower = exp(estimate - 1.96 * std.error)
  ) |>
  select(term, estimate, OR, CI_lower, CI_upper, p.value) |>
  kable(digits = 3, caption = "Proposed Model Summary")
```

### Residuals vs Fitted Values Plot
We examine the residuals of the proposed model to assess its fit.
```{r}
# Add residuals and predictions
proposed_results <- birthweight_clean |> 
  add_residuals(proposed_model) |> 
  add_predictions(proposed_model)

# Plot residuals vs fitted values
ggplot(proposed_results, aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Residuals vs Fitted Values (Proposed Model)",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_minimal()
```

The residual plot showed acceptable dispersion but slight heteroscedasticity at higher fitted values. This indicates the model could improve but adequately predicts across most of the data range.

## Alternative Models

### Model 2: Birth Length and Gestational Age
This model uses *birth length* (`blength`) and *gestational age* (`gaweeks`) as predictors without interactions.

```{r}
# Fit Model 2
model_2 <- lm(bwt ~ blength + gaweeks, data = birthweight_clean)

# Summarize Model 2
model_2_summary <- model_2 |> broom::tidy()
model_2_summary |> 
  kable(digits = 3, caption = "Model 2 Summary")

# Add residuals and predictions
model_2_results <- birthweight_clean |> 
  add_residuals(model_2) |> 
  add_predictions(model_2)

# Plot residuals vs fitted values
ggplot(model_2_results, aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Residuals vs Fitted Values (Model 2)",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_minimal()
```

More dispersion around fitted values than the proposed model, indicating less accurate predictions.

### Model 3: Head Circumference, Length, and Baby’s Sex with Interactions
This model includes interactions between *head circumference* (`bhead`), *length* (`blength`), and *baby sex* (`babysex`).
```{r}
# Fit Model 3
model_3 <- lm(bwt ~ bhead * blength * babysex, data = birthweight_clean)

# Summarize Model 3
model_3_summary <- model_3 |> broom::tidy()
model_3_summary |> 
  kable(digits = 3, caption = "Model 3 Summary")

# Add residuals and predictions
model_3_results <- birthweight_clean |> 
  add_residuals(model_3) |> 
  add_predictions(model_3)

# Plot residuals vs fitted values
ggplot(model_3_results, aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Residuals vs Fitted Values (Model 3)",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_minimal()
```

Residuals showed a tighter cluster, suggesting better model fit compared to Model 2 but not necessarily outperforming the proposed model.


## Cross-Validation
Evaluate all three models using Monte Carlo cross-validation (100 iterations, 80/20 split).
```{r}
# Perform cross-validation
cv_results <- crossv_mc(birthweight_clean, n = 100, test = 0.2) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble),
    proposed_model = map(train, ~ lm(bwt ~ babysex + gaweeks + ppwt + blength + blength:gaweeks, data = .x)),
    model_2 = map(train, ~ lm(bwt ~ blength + gaweeks, data = .x)),
    model_3 = map(train, ~ lm(bwt ~ bhead * blength * babysex, data = .x)),
    rmse_proposed = map2_dbl(proposed_model, test, ~ {
      preds <- predict(.x, newdata = .y)
      sqrt(mean((.y$bwt - preds)^2))
    }),
    rmse_model_2 = map2_dbl(model_2, test, ~ {
      preds <- predict(.x, newdata = .y)
      sqrt(mean((.y$bwt - preds)^2))
    }),
    rmse_model_3 = map2_dbl(model_3, test, ~ {
      preds <- predict(.x, newdata = .y)
      sqrt(mean((.y$bwt - preds)^2))
    })
  )

# Prepare cv_summary for plotting
cv_summary <- cv_results |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    cols = everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) |>
  mutate(model = factor(model, levels = c("proposed", "model_2", "model_3")))

# Plot RMSE comparison
ggplot(cv_summary, aes(x = model, y = rmse, fill = model)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) + 
  geom_jitter(width = 0.2, alpha = 0.5, color = "black") + 
  scale_fill_manual(values = c("lightgreen", "salmon", "skyblue")) + 
  labs(
    title = "Cross-Validation RMSE Comparison",
    x = "Model",
    y = "RMSE",
    fill = "Model"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```

- **Proposed Model:** Median RMSE ~330
- **Model 2:** Median RMSE ~350
- **Model 3:** Median RMSE ~310

The RMSE distribution for each model was visualized using a boxplot:**Model 3** showed the lowest RMSE variation and **Proposed Model** performed consistently better than Model 2.

The proposed model outperformed Model 2 but was slightly less predictive than Model 3, as indicated by the RMSE comparison. The interaction terms in Model 3 likely enhanced its predictive power.









































